---
title: "WEC_TM_3"
author: "Tymoteusz Mętrak"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Libraries
```{r}
library(secr) # for point within the polygon
library(tidyr)
library(tidyverse)
library(sf) # simple features -> do map
library(dlookr) # do przyjemnej EDA
library(ggplot2) # do wykresów
library(ggthemes) # for elegant themes
library(stringr) # do operacji na tekście
library(openxlsx) # do otwierania excela
library(lmtest) # do testów statystycznych
library(car) # i.a. for vif test
library(spatialreg) # for spatial regressions
library(spdep) # for spatial objects
library(spgwr) # for GWR estimation
library(sandwich) # for robust standard errors
```

Reading the data
```{r}
getwd()
dane <- read.xlsx("datamaster_final_end.xlsx")
colnames(dane)[63] <- "type_500k"
dane <- dane %>% 
  select(-teryt_gmina) %>% 
  select(-glosy_wazne)
dane_old <- read.csv("data_municipalities.csv")
```

Other data reading
```{r}
counties <- read.csv("data_counties.csv")

gminy <- read_sf("shapefiles_wec/map_municipalities.shp")
colnames(gminy)[1] <- colnames(dane)[1]
colnames(dane)[1]
gminy$municipality_code <- as.numeric(gminy$municipality_code)

gminy <-  left_join(gminy, dane, by = join_by(municipality_code))

zabory <- read_sf("shapefiles_wec/map_partitions.shp")
```


Adjusting projections
```{r}
gminy <- st_transform(gminy, 4326)
zabory <- st_transform(zabory, 4326)
```

The map
```{r}
ggplot(gminy) +
  geom_sf(aes(fill=percent_vaccinated,col=percent_vaccinated)) +
  geom_sf(data = zabory, fill = NA, col = "red")+
  scale_fill_viridis_c()+
  guides(fill=guide_legend(title = "Vaccination \nrate"),
         color="none")+
  labs(title = "Vaccination rate in Poland",
       x="\n",
       y="\n")+
  theme_gdocs()
```

Weights matrix
```{r}
crds <- st_centroid(gminy)
crds <- st_transform(crds, 2180)
nbl = tri2nb(crds$geometry) #    
cont.listw<-nb2listw(nbl, style="W")
```
first formula
```{r}
my_formula <- as.formula(                      # Create formula
  paste("percent_vaccinated ~ ", paste(colnames(dane)[c(5:ncol(dane))], collapse = " + ")))
my_formula
```

OLS
```{r}
model_OLS <- lm(formula = my_formula, 
             weights = dane_old$population_total,
             data = dane)
summary(model_OLS)
```

```{r}
my_formula2 <- as.formula("percent_vaccinated ~ area_km2 + urbanization_rate +
                               healthcare_advices + installations_central_heating +
                                 persons_per_appartment + persons_per_library +
                               forests_area + 
                                 bicycle_paths_per_10k_persons  +
                               revenues_per_capita_PIT  +
                                 revenues_per_capita_CIT +
                               children_3_5_in_kindergartens + 
                                 birthrate_per_1000_persons +
                                 Prussian + 
                                 Russian + 
                                 frekwencja + 
                                 glosy_KO+
                                 glosy_KONF + 
                                 glosy_PSL + 
                                 glosy_SLD+
                                 high_exposure +
                                 index_ineq +
                                 average_wage_relative +
                                 doctors_per_1000_persons +
                                 cars_per_1000_persons +
                                 education_share_higher +
                                 tourits_per_1000_persons+
                                 rolny_revenue_ratio +
                                 long+
                                 lat")
```

OLS2
```{r}
model_OLS2 <- lm(formula = my_formula2, 
             weights = dane_old$population_total,
             data = dane)
summary(model_OLS2)
```

```{r}
vif(model_OLS2)
```


Breusch-Pagan test for heteroskedasity
```{r}
bptest(model_OLS2)
```
I reject the hypothesis, therefore run robust model



OLS robust errors clustered for county
```{r}
vcov = vcovHC(model_OLS2, type="HC0", cluster=~county_code)
rob <- coeftest(model_OLS2, vcov.=vcov)
rob
```


```{r}
my_formula3 <- as.formula("percent_vaccinated ~ area_km2  +
                               healthcare_advices +
                                 persons_per_appartment  +
                               forests_area + 
                                 bicycle_paths_per_10k_persons  +
                                 revenues_per_capita_PIT  +
                               children_3_5_in_kindergartens + 
                                 birthrate_per_1000_persons +
                                 Prussian + 
                                 Russian + 
                                 frekwencja + 
                                 glosy_KO+
                                 glosy_KONF + 
                                 glosy_PSL + 
                                 glosy_SLD+
                                 high_exposure +
                                 index_ineq +
                                 average_wage_relative +
                                 doctors_per_1000_persons +
                                 cars_per_1000_persons +
                                 education_share_higher +
                                 tourits_per_1000_persons+
                                 rolny_revenue_ratio +
                                 long+
                                 lat")
```

OLS3
```{r}
model_OLS3 <- lm(formula = my_formula3, 
             weights = dane_old$population_total,
             data = dane)

vcov3 = vcovHC(model_OLS3, type="HC0", cluster=~county_code)
rob3 <- coeftest(model_OLS3, vcov.=vcov3)
rob3
```

And compare the models
```{r}
stargazer(model_OLS, rob, type="text")
```


```{r}
crds <- st_centroid(gminy)
crds <- st_transform(crds, 2180)

nbl = tri2nb(crds$geometry) 
cont.listw<-nb2listw(nbl, style="W")

result_moran<-moran.test(dane$percent_vaccinated, cont.listw) 
result_moran
```
The test states that there is a strong spatial dependence (p-value virtually 0). The Moran I statistic is very high and equal to 0.71. We will build spatial model then.


SEM
```{r}
SEM<-errorsarlm(formula = my_formula3, data = dane,
              listw=cont.listw)
summary(SEM)
```

```{r}
stargazer(model_OLS3, rob3,SEM, type="text")
stargazer(model_OLS3, rob3, SEM,
          model.names = F,
          column.labels = c("OLS", "OLS with robust s.e.", "Spatial Error Model"),
          column.sep.width = "1pt",
          dep.var.labels = "Vaccinated population in municipality (%)",
          title = "OLS estimates2",
          type ="html",
          out="OLS_estimates2.html")
```


```{r}
SAR_TSLS <- stsls(formula = my_formula3, data = dane,
                    listw=cont.listw)
W <- as(cont.listw, "CsparseMatrix")
trMatc <- trW(W, type="mult")
loobj1 <- spatialreg::impacts(SAR_TSLS, R=2000, tr=trMatc)
(sums <- summary(loobj1, zstats=TRUE, short=TRUE))
```

Do tabeli wyników
```{r}
STLS_df <- data.frame(sums$res)
```

Oszacowane, ale niewykorzystane - mam TSLS
```{r}
SAR <- lagsarlm (formula = my_formula3, data = dane,
                 listw=cont.listw)
stargazer(SAR, type="text")
```


```{r}
Prussia <- dane[dane$Prussian==1,]
Russia <- dane[dane$Russian==1,]
Austria <- dane[dane$Prussian ==0 & dane$Russian==0,]

dane_old2 <- data.frame(dane_old, dane$Prussian, dane$Russian)
dane_old_Prussia <- dane_old2[dane_old2$dane.Prussian==1,]
dane_old_Russia <- dane_old2[dane_old2$dane.Russian==1,]
dane_old_Austria <- dane_old2[dane_old2$dane.Prussian==0 & dane_old2$dane.Russian==0,]
```

```{r}
my_formula4 <- as.formula("percent_vaccinated ~ area_km2  +
                               healthcare_advices +
                                 persons_per_appartment  +
                               forests_area + 
                                 bicycle_paths_per_10k_persons  +
                                 revenues_per_capita_PIT  +
                               children_3_5_in_kindergartens + 
                                 birthrate_per_1000_persons +
                                 frekwencja + 
                                 glosy_KO+
                                 glosy_KONF + 
                                 glosy_PSL + 
                                 glosy_SLD+
                                 high_exposure +
                                 index_ineq +
                                 average_wage_relative +
                                 doctors_per_1000_persons +
                                 cars_per_1000_persons +
                                 education_share_higher +
                                 tourits_per_1000_persons+
                                 rolny_revenue_ratio +
                                 long+
                                 lat")
```


```{r}
model_Prussia <- lm(formula = my_formula4, 
             weights = dane_old_Prussia$population_total,
             data = Prussia)

vcov_Prussia = vcovHC(model_Prussia, type="HC0", cluster=~county_code)
rob_Prussia <- coeftest(model_OLS3, vcov.=vcov_Prussia)

model_Russia <- lm(formula = my_formula4, 
             weights = dane_old_Russia$population_total,
             data = Russia)

vcov_Russia = vcovHC(model_Russia, type="HC0", cluster=~county_code)
rob_Russia <- coeftest(model_OLS3, vcov.=vcov_Russia)

model_Austria <- lm(formula = my_formula4, 
             weights = dane_old_Austria$population_total,
             data = Austria)

vcov_Austria = vcovHC(model_Austria, type="HC0", cluster=~county_code)
rob_Austria <- coeftest(model_OLS3, vcov.=vcov_Austria)

stargazer(rob_Prussia, rob_Russia, rob_Austria, type="text")
```




